---
title: "04 Active-site conservation"
output: html_notebook
---

```{r}
library(ggplot2)
library(data.table)
library(scales)
library(patchwork)
```

# Setup

```{r}
load("source_variables.RData")

output_dir_fig <- paste0(results_path, "Fig3/")
output_dir_figS5 <- paste0(results_path, "FigS5/")
output_dir_chimera <- paste0(results_path, "00_chimera/")
dir.create(output_dir_fig, showWarnings = FALSE, recursive = TRUE)
dir.create(output_dir_figS5, showWarnings = FALSE, recursive = TRUE)

```

# Read ddG

```{r}
all_ddg<-read.csv(paste0(results_newdata,"all_ddg_norm.csv"))

setDT(all_ddg)
setnames(all_ddg,"mean_kcal.mol","mean_kcal/mol_old")
setnames(all_ddg,"std_kcal.mol","std_kcal.mol_old")
setnames(all_ddg,"ddg_norm","mean_kcal/mol")
setnames(all_ddg,"std_kcal_norm","std_kcal.mol")

all_ddg[,mut_aa:=substr(as.character(id),nchar(as.character(id)),nchar(as.character(id)))]

all_ddg<-all_ddg[mut_aa!="*"]

all_ddg_a<-all_ddg[assay=="Activity",]
fwrite(all_ddg,paste0(results_newdata,"all_ddg_norm_clean.csv"))
```

# Measurement counts and active-site size

```{r manuscript_ddg_counts_and_as_size}
cat("=== ddG measurement counts (non-WT, non-stop; both assays) ===\n")
print(all_ddg[, .(n_measurements = .N), by = assay])
cat(sprintf("ddGf (Folding)  measurements: %d\n", all_ddg[assay == "Folding",  .N]))
cat(sprintf("ddGa (Activity) measurements: %d\n", all_ddg[assay == "Activity", .N]))

cat("\n=== Active-site size (KinCore positions with active_site != \"\") ===\n")
print(all_ddg_a[active_site != "", .(n_positions = uniqueN(Pos_al)), by = active_site])
cat(sprintf("Total active-site positions: %d\n",
            uniqueN(all_ddg_a[active_site != "", Pos_al])))
cat(sprintf("Active-site ddGa measurements: %d\n", all_ddg_a[active_site != "", .N]))
```

# Active-site region medians

```{r manuscript_region_medians}
region_median <- function(label, mask) {
  v <- all_ddg_a[mask, `mean_kcal/mol`]
  cat(sprintf("%-64s median ddGa = %.2f (n = %d mutations)\n",
              label, median(v, na.rm = TRUE), length(v)))
}
cat("=== Active-site region median ddGa (median of mutation ddGa) ===\n")
region_median("Mg-positioning loop (active_site == 'Mg positioning')",
              all_ddg_a$active_site == "Mg positioning")
region_median("Catalytic-loop ATP contacts (aligned 161,163,165,166,168)",
              all_ddg_a$Pos_al %in% c(161, 163, 165, 166, 168))
region_median("N-terminal activation-loop substrate segment (aligned 202-206)",
              all_ddg_a$Pos_al %in% 202:206)
region_median("Glycine-rich loop (aligned 34-39)",
              all_ddg_a$Pos_al %in% 34:39)
```


# Per-position median ddGa

```{r}
ddg_median_a <- all_ddg_a[, .(pos_median = median(`mean_kcal/mol`, na.rm = TRUE)),
                                 by = .(Pos, kinase, assay,Pos_al)]
all_ddg_a <- merge(all_ddg_a, ddg_median_a,
                   by = c("Pos", "kinase", "assay","Pos_al"),
                   all.x = TRUE)
```

# Activity classifier

```{r}
all_ddg_a[,z_twotailed:=(abs(`mean_kcal/mol`)-0.2)/`std_kcal.mol`]
all_ddg_a[,p_twotailed:=1-pnorm(z_twotailed)]
all_ddg_a[,fdr_twotailed:=p.adjust(p_twotailed,method = "fdr")]


all_ddg_a[,class:="no change"]
all_ddg_a[fdr_twotailed<0.1 & `mean_kcal/mol`< -0.2,class:="activating"]
all_ddg_a[fdr_twotailed<0.1 & `mean_kcal/mol`>0.2,class:="inactivating"]

cat("=== Activating and inactivating mutations ===\n")
print(all_ddg_a[class != "no change", .N, by = class])
print(dcast(all_ddg_a[class != "no change", .N, by = .(kinase, class)], kinase ~ class, value.var = "N"))
```


# Cross-kinase divergence
```{r}
active_site_data<-all_ddg_a[active_site!=""]
active_site_data_pos_only<-unique(active_site_data$Pos_al)
active_site_data_pos<-unique(all_ddg_a[active_site!="",.(Pos_al,active_site)])

all_active_site_data<-all_ddg_a[Pos_al%in%c(active_site_data_pos_only)]

mut_divergence <- all_active_site_data[, {
  vals <- `mean_kcal/mol`[!is.na(`mean_kcal/mol`)]
  list(mut_div = if (length(vals) >= 2) mean(dist(vals)) else NA_real_)
}, by = .(Pos_al, mut_aa)]

global_pos_divergence <- mut_divergence[, .(
  pos_divergence = mean(mut_div, na.rm = TRUE)
), by = Pos_al]
global_pos_divergence <- merge(global_pos_divergence, active_site_data_pos)

p_divergence_strip <- ggplot(global_pos_divergence, aes(x = as.factor(Pos_al), y = "Divergence", fill = pos_divergence)) +

	geom_tile() +

  scale_fill_gradient2(
  low = "darkorange4",
  mid = "orange",
  high = "grey95",
  midpoint = 0.35,
  limits = c(0.2, 0.5),
  oob = scales::squish,
  name = "Mean\nDiff" )+

	theme_classic(base_size = 14) +
	labs(
		x = "",
		y = ""
	) +
	theme(
		axis.text.y = element_blank(),
		axis.ticks.y = element_blank(),
		panel.grid = element_blank(),
		legend.position = "right",
		legend.title = element_text(size = 15),
		legend.text = element_text(size = 15),
		legend.key.height = unit(0.4, "cm"),
		legend.key.width  = unit(0.6, "cm"),
		axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1),
		strip.text =  element_blank()

	) +

	scale_x_discrete(expand = c(0, 0)) +
	scale_y_discrete(expand = c(0, 0))+
  facet_grid(~active_site, scales = "free_x",space = "free_x",
             labeller = labeller(active_site = label_wrap_gen(width = 10)))

```

# Adenosine region

```{r manuscript_adenosine_region}
aden_pos <- c(88, 110, 111, 112, 113, 117)

v_aden <- all_ddg_a[Pos_al %in% aden_pos, `mean_kcal/mol`]
cat("=== Adenosine region ===\n")
cat(sprintf("Pooled median ddGa over adenosine residues (%s): %.2f (n = %d mutations)\n",
            paste(aden_pos, collapse = ","), median(v_aden, na.rm = TRUE), length(v_aden)))

atp_pos  <- unique(global_pos_divergence[active_site == "ATP contact", Pos_al])
aden_div <- global_pos_divergence[Pos_al %in% aden_pos, pos_divergence]
rest_div <- global_pos_divergence[Pos_al %in% setdiff(atp_pos, aden_pos), pos_divergence]
cat(sprintf("Divergence, adenosine residues:          mean = %.2f  median = %.2f  (n = %d)\n",
            mean(aden_div), median(aden_div), length(aden_div)))
cat(sprintf("Divergence, rest of ATP-contact residues: mean = %.2f  median = %.2f  (n = %d)\n",
            mean(rest_div), median(rest_div), length(rest_div)))
```

# Active-site heatmap
```{r}
cap_val <- quantile(abs(active_site_data$pos_median), probs = 0.99, na.rm = TRUE)
heatmap_plot_dt<-unique(active_site_data[,.(Pos_al,kinase,pos_median,ATP,active_site,WT_AA)])

p_heatmap_dist_activesite <-ggplot(heatmap_plot_dt,
                                    aes(x = as.factor(Pos_al),
                                        y = factor(kinase, levels=rev(kinases)),
                                        fill = pos_median)) +
  geom_tile(color = "white", size = 0.3) +
geom_text(
  data = heatmap_plot_dt[ATP=="ATP contact"],
  aes(x = as.factor(Pos_al), y = kinase, label = WT_AA),
  color = "#ffe500", size = 5,fontface = "bold"
) +
geom_text(
  data = heatmap_plot_dt[ATP == "no ATP contact"&active_site!=""],
  aes(x = as.factor(Pos_al), y = kinase, label = WT_AA),
  color = "black", size = 5
)+
  scale_fill_gradientn(
    colours = c("blue3", "grey90", "brown3"),
    values = scales::rescale(c(-cap_val, 0, cap_val)),
    limits = c(-cap_val, cap_val),
    guide = "colorbar",
    oob = scales::squish
  ) +
  theme_classic() +
  labs(x = "", y = "", fill =expression(Delta*Delta*G[a]), color = "") +
  theme(
    axis.ticks = element_blank(),
    axis.ticks.x.top = element_line(),
    plot.title = element_blank(),
    legend.title = element_text(size = 15, face = "bold"),
    legend.text = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title = element_text(size = 11),
    axis.text.x = element_blank(),

    strip.text =  element_text(size = 15, face = "bold")
  ) +
    facet_grid(~active_site, scales = "free_x",space = "free_x",
             labeller = labeller(active_site = label_wrap_gen(width = 10)))+
  scale_y_discrete(expand = expansion(mult = c(0, 0.25)), labels = relabel_jnk2)+
      guides(pattern = "none", color = "none")


```
## Median across kinases


```{r}
global_pos_effect <- all_active_site_data[, .(
	pos_median_all = median(pos_median, na.rm = TRUE)
), by = Pos_al]
global_pos_effect<-merge(global_pos_effect,unique(active_site_data[,.(Pos_al,Pos,pos_median,active_site)]))

p_effect_strip <- ggplot(global_pos_effect, aes(x = as.factor(Pos_al), y = "Median", fill = pos_median_all)) +

	geom_tile() +

  scale_fill_gradientn(
    colours = c("blue3", "grey90", "brown3"),
    values = scales::rescale(c(-cap_val, 0, cap_val)),
    limits = c(-cap_val, cap_val),
    guide = "colorbar",
    oob = scales::squish
  ) +

	theme_classic(base_size = 14) +
	labs(
		x = "",
		y=""
	) +
	theme(
	  axis.text.y = element_text(size = 15,color="black"),
		axis.text.x = element_blank(),
		axis.ticks = element_blank(),
		panel.grid = element_blank(),
		legend.position = "none",
		   strip.text =  element_blank()

	) +

	scale_x_discrete(expand = c(0, 0)) +
	scale_y_discrete(expand = c(0, 0))+
  facet_grid(~active_site, scales = "free_x",space = "free_x",
             labeller = labeller(active_site = label_wrap_gen(width = 10)))


p_heatmap_conserv_AS<-p_heatmap_dist_activesite/p_effect_strip/p_divergence_strip+ plot_layout(heights = c(5,0.6, 0.6))

```


## Save

```{r}
ggsave(p_heatmap_conserv_AS,filename =paste0(output_dir_fig,"Fig3A_active_site_heatmap",format_img),height = 5,width = 15,limitsize = FALSE)
```


# ChimeraX

## Cross-kinase divergence

```{r}
pos_global_pos_divergence<-unique(merge(all_active_site_data[,.(Pos_al,kinase,Pos)],global_pos_divergence, all.x=TRUE,by="Pos_al"))


for (k in kinases) {
  sig_residues <- pos_global_pos_divergence[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: ASconservation",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$pos_divergence)
  )

  file_name <- paste0(output_dir_chimera, "AS_conservation_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```

## Median effect across kinases

```{r}
pos_global_pos_effect<-unique(merge(all_active_site_data[,.(Pos_al,kinase,Pos)],global_pos_effect, all.x=TRUE, by=c("Pos_al","Pos")))


for (k in kinases) {
  sig_residues <- pos_global_pos_effect[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: ASeffect",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$pos_median_all)
  )

  file_name <- paste0(output_dir_chimera, "AS_avg_effect_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```

## Median effect per kinase

```{r}
pos_effect<-unique(all_active_site_data[,.(Pos_al,kinase,Pos,pos_median)])


for (k in kinases) {
  sig_residues <- pos_effect[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: ASeffect",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$pos_median)
  )

  file_name <- paste0(output_dir_chimera, "AS_avg_effect_each_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```

## Median ddGa per position

```{r}
ddg_median_a_unique <- unique(ddg_median_a[assay == "Activity", .(Pos, kinase, pos_median)])

for (k in kinases) {
  sig_residues <- ddg_median_a_unique[kinase == k]
  if (nrow(sig_residues) == 0) next
  writeLines(c(
    "attribute: DDGmedian",
    "recipient: residues",
    paste0("\t:", sig_residues$Pos, "\t", sig_residues$pos_median)
  ), paste0(output_dir_chimera, "DDGa_median_", k, ".defattr"))
}
```

# Active-site enrichment

```{r active_site_enrichment}
as_positions <- unique(all_ddg_a[active_site != "", Pos_al])
all_ddg_a[, in_AS := Pos_al %in% as_positions]

as_enrich_one <- function(dt, cls) {
  a <- nrow(dt[in_AS == TRUE  & class == cls])
  b <- nrow(dt[in_AS == TRUE  & class != cls])
  cc <- nrow(dt[in_AS == FALSE & class == cls])
  d <- nrow(dt[in_AS == FALSE & class != cls])
  ft   <- fisher.test(matrix(c(a, b, cc, d), nrow = 2))
  or_h <- ((a + 0.5) * (d + 0.5)) / ((b + 0.5) * (cc + 0.5))
  se   <- sqrt(1/(a + 0.5) + 1/(b + 0.5) + 1/(cc + 0.5) + 1/(d + 0.5))
  data.table(n_class_AS = a, n_AS = a + b, n_class_total = a + cc,
             OR = unname(ft$estimate), p = ft$p.value,
             OR_plot = or_h,
             ci_lo = exp(log(or_h) - 1.96 * se),
             ci_hi = exp(log(or_h) + 1.96 * se))
}

as_enrich <- rbindlist(lapply(c("inactivating", "activating"), function(cls) {
  per    <- rbindlist(lapply(kinases, function(k)
              cbind(kinase = k, class = cls, as_enrich_one(all_ddg_a[kinase == k], cls))))
  pooled <- cbind(kinase = "All", class = cls, as_enrich_one(all_ddg_a, cls))
  rbind(per, pooled)
}))
as_enrich[, fdr := p.adjust(p, method = "fdr")]

print(as_enrich[, .(kinase, class, n_class_AS, n_AS, n_class_total,
                    OR = round(OR, 2), p = signif(p, 2), fdr = signif(fdr, 2))])
```

## Plot

```{r active_site_enrichment_fig}
pd <- copy(as_enrich)
pd[, kinase := factor(kinase, levels = c(as.character(kinases), "All"))]
pd[, class  := factor(class, levels = c("inactivating", "activating"))]

pd[, log2_OR := log2(OR_plot)]
pd[, sig := fcase(fdr < 1e-5, "***", fdr < 1e-3, "**", fdr < 0.1, "*", default = "")]
pd[, star_y := ifelse(log2_OR >= 0, log2_OR + 0.25, log2_OR - 0.45)]

dodge <- position_dodge(width = 0.7)
p_as_enrich <- ggplot(pd, aes(x = kinase, y = log2_OR, fill = class)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_col(position = dodge, width = 0.65, colour = "grey20", linewidth = 0.3) +
  geom_text(aes(y = star_y, label = sig), position = dodge,
            size = 5, vjust = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c(inactivating = "brown3", activating = "blue3"),
                    name = NULL) +
  labs(x = NULL, y = "Enrichment in active site (Log2 OR)") +
  scale_x_discrete(labels = relabel_jnk2) +
  theme_classic(base_size = 14) +
  theme(legend.position = "top",
        axis.text.x = element_text(color = "black"))
ggsave(p_as_enrich,
       filename = paste0(output_dir_figS5, "FigS5B_active_site_enrichment", format_img),
       height = 3.3, width = 4.5, limitsize = FALSE)
```

# Activating mutations at adenosine-contact positions

```{r adenosine_ring_perposition_or}
aden_perpos_or <- rbindlist(lapply(aden_pos, function(pos) {
  rbindlist(lapply(kinases, function(k) {
    dt <- all_ddg_a[kinase == k & Pos_al %in% as_positions]
    A <- nrow(dt[Pos_al == pos & class == "activating"])
    B <- nrow(dt[Pos_al == pos & class != "activating"])
    C <- nrow(dt[Pos_al != pos & class == "activating"])
    D <- nrow(dt[Pos_al != pos & class != "activating"])
    ft <- fisher.test(matrix(c(A, B, C, D), nrow = 2))
    data.table(Pos_al = pos, kinase = k, WT_AA = dt[Pos_al == pos, WT_AA][1],
               n_act = A, n = A + B, OR = unname(ft$estimate), p = ft$p.value)
  }))
}))
aden_perpos_or[, kinase := factor(kinase, levels = kinases)]
setorder(aden_perpos_or, Pos_al, kinase)
print(aden_perpos_or[, .(Pos_al, kinase, WT_AA, n_act, n, OR = round(OR, 2), p = signif(p, 2))])
```

