---
title: "05 Distance decay"
output: html_notebook
---

```{r}
library(ggplot2)
library(data.table)
library(scales)
library(patchwork)
library(ggpointdensity)
```

# Setup
```{r}
load("source_variables.RData")

output_dir_fig <- paste0(results_path, "Fig4/")
dir.create(output_dir_fig, showWarnings = FALSE, recursive = TRUE)
output_dir_figS <- paste0(results_path, "FigS6/")
dir.create(output_dir_figS, showWarnings = FALSE, recursive = TRUE)
output_dir_chimera <- paste0(results_path, "00_chimera/")
```


# Read ddG

```{r}
all_ddg<-fread(paste0(results_newdata,"all_ddg_norm_clean.csv"))


all_ddg_a<-all_ddg[assay=="Activity"]
```

# Per-position ddGa summaries

```{r}
ddg_mean_a <- all_ddg_a[, .(pos_mean = weighted.mean(`mean_kcal/mol`, `std_kcal.mol`^-2, na.rm = TRUE)),
                                by = .(Pos, kinase, assay,Pos_al)]
all_ddg_a <- merge(all_ddg_a, ddg_mean_a,
                  by = c("Pos", "kinase", "assay","Pos_al"),
                  all.x = TRUE)
ddg_median <- all_ddg[, .(pos_median = median(`mean_kcal/mol`, na.rm = TRUE)),
                                 by = .(Pos, kinase, assay,Pos_al)]
all_ddg_a <- merge(all_ddg_a, ddg_median,
                   by = c("Pos", "kinase", "assay","Pos_al"),
                   all.x = TRUE)
```

# Distance to the active site
```{r}
dist_ATP <- fread(file.path(data_struct, "minimum_residuedistances_ATP.txt"))

anno_dt <- copy(all_ddg_a)
anno_dt <- as.data.table(anno_dt)


ATP_contacts <- unique(anno_dt[
  ATP == "ATP contact",
  .(kinase, Pos, Pos_al, ATP)
])

anno_dt[, Pos := as.numeric(Pos)]
dist_ATP[, Pos := as.numeric(Pos)]

all_ddg_dist <- merge(anno_dt,dist_ATP, by = c("kinase", "Pos"),all.x=TRUE)
```

# Contact shells

## Second shell

```{r}
dist_dt <- fread(paste0(data_struct, "minimum_residuedistances_pairwise.txt"))

dist_dt2 <- copy(dist_dt)
setnames(dist_dt2, c("Pos1", "Pos2"), c("Pos2", "Pos1"))


dist_dt2[, is_atp := Pos1 %in% ATP_contacts[kinase == .BY$kinase, Pos],
         by = kinase]

dist_dt2[, Pos := fifelse(is_atp, Pos2, Pos1)]
dist_dt2 <- dist_dt2[is_atp == TRUE]


dist_closest <- dist_dt2[
  , .SD[which.min(scHA_ca)], by = .(kinase, Pos)
]
dist_closest_clean<-dist_closest[,.(kinase,Pos,scHA_ca)]
dist_closest_clean[, is_atp := Pos %in% ATP_contacts[kinase == .BY$kinase, Pos],
         by = kinase]
second_shell <- dist_closest_clean[
  scHA_ca < 5&is_atp==FALSE,
]

all_ddg_dist[, second_shell := Pos %in% second_shell[kinase == .BY$kinase, Pos], by = kinase]
```

## Shell assignment

```{r}
distance_cutoff <- 5

shells <- list()
shells[[1]] <- unique(ATP_contacts[, .(kinase, Pos)])

repeat {
  prev_shell <- shells[[length(shells)]]

  close_res <- dist_dt[
    prev_shell,
    on = .(kinase, Pos1 = Pos),
    nomatch = 0
  ][scHA_ca < distance_cutoff, .(kinase, Pos = Pos2)]

  already <- unique(rbindlist(shells, use.names = TRUE))
  close_res <- close_res[!already, on = .(kinase, Pos)]

  if (nrow(close_res) == 0) break

  shells[[length(shells) + 1]] <- unique(close_res)
}

shell_dt <- rbindlist(lapply(seq_along(shells), \(i) {
  shells[[i]][, shell := i]
}), use.names = TRUE)

all_ddg_dist[
  , shell := shell_dt[.SD, on = .(kinase, Pos), x.shell]
]


```

## Shell spacing

```{r shell-spacing}
shell_dist_ladder <- all_ddg_dist[is.finite(active_site_distance) & !is.na(shell),
                                  .(mean_dist = mean(active_site_distance)), by = shell]
setorder(shell_dist_ladder, shell)
shell_spacing <- mean(diff(shell_dist_ladder$mean_dist))
print(shell_dist_ladder)
cat(sprintf("Mean spacing between consecutive shells = %.1f A\n", shell_spacing))
```

## ChimeraX: shells

```{r}
shells<-unique(all_ddg_dist[,.(kinase,Pos,shell)])
for (k in kinases) {
  sig_residues <- shells[kinase == k]
  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: Shells",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$shell)
  )

  file_name <- paste0(output_dir_chimera, "shells_", k, ".defattr")


  writeLines(chimerax_lines, file_name)
}
```


# Activity classifier

```{r}
all_ddg_dist[,z_twotailed:=(abs(`mean_kcal/mol`)-0.2)/`std_kcal.mol`]
all_ddg_dist[,p_twotailed:=1-pnorm(z_twotailed)]
all_ddg_dist[,fdr_twotailed:=p.adjust(p_twotailed,method = "fdr")]


all_ddg_dist[,class:="no change"]
all_ddg_dist[fdr_twotailed<0.1 & `mean_kcal/mol`< -0.2,class:="activating"]
all_ddg_dist[fdr_twotailed<0.1 & `mean_kcal/mol`>0.2,class:="inactivating"]
```


# Distance decay

## Absolute ddGa
```{r}
all_ddg_dist[,abs_ddg:=abs(`mean_kcal/mol`)]
all_ddg_dist[, kinase := factor(kinase, levels =kinases)]
```

## Fit functions
```{r}
fit_exponential_curve <- function(xvector, yvector, tit, plotfig = TRUE, color_vector = NULL, point_size = 1.5, anno_y = 2, anno_size = 3.88, anno_gap = NULL, err_vector = NULL) {
  data <- data.table(x = xvector, y = yvector)
  if (!is.null(color_vector)) data[, color_group := color_vector]
  if (!is.null(err_vector))   data[, err := err_vector]

  residual_sum_of_squares <- function(params, x, y) {
    a <- params[1]
    b <- params[2]
    predicted <- a * exp(b * x)
    sum((y - predicted)^2)
  }

  initial_guess <- c(a = 2.5, b = -0.1)
  optimized_params <- optim(initial_guess, residual_sum_of_squares, x = data$x, y = data$y)$par

  fit <- nls(y ~ a * exp(b * x), data = data,
             start = list(a = optimized_params[1], b = optimized_params[2]))

  data[, predicted := predict(fit, newdata = .SD)]
  data[, residual := y - predicted]

  plot_obj <- NULL
  if (plotfig) {
    x_seq <- seq(min(data$x), max(data$x), length.out = 100)
    fitted_values <- predict(fit, newdata = data.frame(x = x_seq))
    parameters <- fit$m$getPars()
    d_half <- log(2) / abs(unname(parameters[2]))
    a0_anno    <- bquote("|" * Delta * Delta * "Ga|" [0] == .(round(unname(parameters[1]), 3)))
    k_anno     <- bquote(k == .(round(unname(parameters[2]), 3)))
    dhalf_anno <- bquote(d[1/2] == .(round(d_half, 1)) * " Å")
    if (is.null(anno_gap)) {
      parameters_anno <- bquote(atop(atop(.(a0_anno), .(k_anno)), .(dhalf_anno)))
      anno_layer <- annotate("text", x = 25, y = anno_y,
                             label = paste(deparse(parameters_anno), collapse = ""),
                             color = "black", parse = TRUE, size = anno_size)
    } else {
      anno_layer <- list(
        annotate("text", x = 25, y = anno_y,                label = paste(deparse(a0_anno),    collapse = ""), color = "black", parse = TRUE, size = anno_size),
        annotate("text", x = 25, y = anno_y - anno_gap,     label = paste(deparse(k_anno),     collapse = ""), color = "black", parse = TRUE, size = anno_size),
        annotate("text", x = 25, y = anno_y - 2 * anno_gap, label = paste(deparse(dhalf_anno), collapse = ""), color = "black", parse = TRUE, size = anno_size)
      )
    }

  err_layer <- if (!is.null(err_vector)) {
    geom_linerange(data = data,
                   aes(x = x, ymin = pmax(0, y - err), ymax = y + err),
                   color = "grey65", alpha = 0.6, linewidth = 0.3)
  } else NULL

plot_obj <- ggplot() +
  err_layer +
  geom_point(
    data = data[color_group != "ATP contact"],
    aes(x = x, y = y),
    color = alpha("grey40", 0.5),
    alpha = 0.6, size = point_size
  ) +
  geom_point(
    data = data[color_group == "ATP contact"],
    aes(x = x, y = y),
    color = "#ffe500",
    alpha = 0.4, size = point_size
  ) +
  geom_line(
    data = data.frame(x = x_seq, Fitted = fitted_values),
    aes(x = x, y = Fitted),
    color = "black", size = 1
  ) +
  anno_layer +
  labs(x = "Distance to the active site (Å)", y =expression("|"*Delta*Delta*"Ga| (kcal/mol)"), title = tit) +
  coord_cartesian(ylim = c(0, 2.5), xlim = c(3, 35)) +
  theme_classic(base_size =13 ) +
  scale_x_continuous(breaks = seq(5, 35, by = 5)) +
  theme(legend.position = "none", axis.text = element_text(size = 11.4))

  }

  message(paste("R² =", round(cor(data$predicted, data$y)^2, 3)))
  return(list(fit = fit, plot = plot_obj, data = data))
}

fit_exponential_curve_for_kinase <- function(kinase_name,dataused, assay_type = "Activity", ycol = "abs_ddg", plotfig = TRUE, point_size = 1.5, anno_y = 2, anno_size = 3.88, anno_gap = NULL) {
  subset_data <- dataused[kinase == kinase_name & assay == assay_type]

  if (nrow(subset_data) < 5) {
    message(paste("Not enough data for kinase:", kinase_name))
    return(NULL)
  }

  xvector <- subset_data$active_site_distance
  yvector <- subset_data[[ycol]]
  color_vector <- ifelse(subset_data$ATP !="ATP contact",ifelse(subset_data$second_shell==TRUE,"second shell", "no ATP contact"), "ATP contact")
  tit <- relabel_jnk2(kinase_name)
  err_vector <- if ("err" %in% names(subset_data)) subset_data$err else NULL

  result <- fit_exponential_curve(xvector, yvector, tit, plotfig, color_vector, point_size, anno_y, anno_size, anno_gap, err_vector = err_vector)

  subset_data[, predicted_y := result$data$predicted]
  subset_data[, residual := result$data$residual]
  subset_data[, color_group := color_vector]

  return(list(fit = result$fit, plot = result$plot, data = subset_data))
}

fit_multiple_kinases <- function(kinases,dataused, assay_type = "Activity", ycol = "abs_ddg", plotfig = TRUE, point_size = 1.5, anno_y = 2, anno_size = 3.88, anno_gap = NULL) {
  results <- list()
  plots <- list()
  pred_data <- list()

  for (k in kinases) {
    message(paste("Fitting:", k))
    res <- fit_exponential_curve_for_kinase(k,dataused, assay_type, ycol, plotfig, point_size, anno_y, anno_size, anno_gap)
    if (!is.null(res)) {
      results[[k]] <- res$fit
      plots[[k]] <- res$plot
      pred_data[[k]] <- res$data
    }
  }

  return(list(fits = results, plots = plots, data = pred_data))
}

```

## All mutations
```{r}
all_ddg_dist$kinase<-factor(all_ddg_dist$kinase,levels=kinases,)
kinase_list <- kinases
results <- fit_multiple_kinases(kinase_list,all_ddg_dist, ycol = "abs_ddg")

all_predicted_data <- rbindlist(results$data, idcol = "kinase")
```

## Per-residue median
```{r}
per_residue_ddg <- all_ddg_dist[, .(
  abs_ddg      = median(abs_ddg, na.rm = TRUE),
  err          = 1.2533 * sd(abs_ddg, na.rm = TRUE) / sqrt(sum(!is.na(abs_ddg))),
  active_site_distance = active_site_distance[1],
  ATP          = ATP[1],
  second_shell = second_shell[1]
), by = .(Pos, kinase, assay, Pos_al)]

per_residue_ddg[, kinase := factor(kinase, levels = kinases)]

results_median <- fit_multiple_kinases(kinases, per_residue_ddg, ycol = "abs_ddg", point_size = 1, anno_y = 1.2, anno_size = 4.9, anno_gap = 0.12)

results_median$plots <- lapply(results_median$plots, function(p)
  p + labs(y = expression("Per residue median |"*Delta*Delta*G[a]*"|")))
p_exp_decay_median <- wrap_plots(results_median$plots, ncol = 5) & coord_cartesian(ylim = c(0, 1.6), xlim = c(3, 35))

ggsave(p_exp_decay_median,
       filename = paste0(output_dir_figS, "FigS6A_decay_median", format_img),
       height = 3.5, width = 17, limitsize = FALSE)
```

# Decay by secondary structure

```{r ss-region-decay}
ss_data <- rbindlist(results_median$data)
reg_map <- unique(all_ddg_a[, .(kinase, Pos_al, Region_Name)])
ss_data <- merge(ss_data, reg_map, by = c("kinase", "Pos_al"), all.x = TRUE)
ss_data <- ss_data[!is.na(Region_Name) & Region_Name != "" &
                   is.finite(active_site_distance) & is.finite(abs_ddg)]
setnames(ss_data, c("abs_ddg", "active_site_distance", "residual"),
                  c("abs_ddga", "dist_as", "resid"))
ss_data[, region_class := fcase(grepl("^H", Region_Name), "Helix",
                                grepl("^B", Region_Name), "Beta strand",
                                rep(TRUE, .N), "Loop")]
ss_data[, region_class := factor(region_class, levels = c("Helix", "Beta strand", "Loop"))]
ss_data[, kinase := factor(kinase, levels = kinases)]

fit_region <- function(d) {
  if (nrow(d) < 10) return(data.table(a=NA_real_, k=NA_real_, k_se=NA_real_, d_half=NA_real_, r2=NA_real_, n=nrow(d)))
  fit <- tryCatch(nls(abs_ddga ~ a * exp(b * dist_as), d, start = list(a = 0.5, b = -0.05),
                      control = nls.control(maxiter = 200, warnOnly = TRUE)), error = function(e) NULL)
  if (is.null(fit)) return(data.table(a=NA_real_, k=NA_real_, k_se=NA_real_, d_half=NA_real_, r2=NA_real_, n=nrow(d)))
  co <- coef(fit); se <- tryCatch(summary(fit)$coefficients["b","Std. Error"], error = function(e) NA_real_)
  data.table(a = co["a"], k = co["b"], k_se = se, d_half = log(2)/abs(co["b"]),
             r2 = cor(fitted(fit), d$abs_ddga)^2, n = nrow(d))
}
region_decay_kin    <- ss_data[, fit_region(.SD), by = .(kinase, region_class)]

add_dhalf_ci <- function(dt) {
  dt[, ci_lo := log(2) / (abs(k) + 1.96 * k_se)]
  dt[, ci_hi := log(2) / (abs(k) - 1.96 * k_se)]
  dt[]
}
region_decay_kin    <- add_dhalf_ci(region_decay_kin)
region_decay_kin[, kinase := factor(kinase, levels = kinases)]
setorder(region_decay_kin, kinase, region_class)
cat("== per-kinase decay by secondary structure: d_half with 95% CI (Wald on k) ==\n")
print(region_decay_kin[region_class %in% c("Beta strand","Helix"),
      .(kinase, region_class, d_half = round(d_half, 1),
        CI = sprintf("%.1f-%.1f", ci_lo, ci_hi), r2 = round(r2, 3), n)])

region_decay_compare <- region_decay_kin[region_class %in% c("Beta strand","Helix"), {
  b <- .SD[region_class == "Beta strand"]; h <- .SD[region_class == "Helix"]
  z <- (abs(b$k) - abs(h$k)) / sqrt(b$k_se^2 + h$k_se^2)
  .(dhalf_beta = round(b$d_half, 1), dhalf_helix = round(h$d_half, 1),
    z = round(z, 2), p_two_sided = signif(2 * pnorm(-abs(z)), 3))
}, by = kinase]
setorder(region_decay_compare, kinase)
cat("\n== beta-strand vs alpha-helix decay constant, z-test (per kinase) ==\n")
print(region_decay_compare)

region_colors    <- c("Helix" = "tomato", "Beta strand" = "lightblue3", "Loop" = "grey50")
region_labels_ss <- expression(alpha*"-helix", beta*"-strand", "Loop")

region_curves_kin <- rbindlist(lapply(seq_len(nrow(region_decay_kin)), function(i) {
  r <- region_decay_kin[i]; if (is.na(r$k)) return(NULL); x <- seq(0, 35, length.out = 100)
  data.table(kinase = r$kinase, region_class = r$region_class, dist_as = x, fitted = r$a * exp(r$k * x)) }))
region_curves_kin[, kinase := factor(kinase, levels = kinases)]
region_curves_kin[, region_class := factor(region_class, levels = levels(ss_data$region_class))]

ann_decay_kin <- copy(region_decay_kin)[!is.na(k)]
ann_decay_kin[, region_class := factor(region_class, levels = levels(ss_data$region_class))]
ann_decay_kin[, kinase := factor(kinase, levels = kinases)]
ann_decay_kin[, y := c("Helix" = 1.80, "Beta strand" = 1.6, "Loop" = 1.4)[as.character(region_class)]]
ann_decay_kin[, lbl := sprintf('k == %.3f * "," ~ d[1/2] == %.1f ~ "Å"', k, d_half)]

p_region_kin <- ggplot() +
  geom_point(data = ss_data, aes(dist_as, abs_ddga, color = region_class), alpha = 0.4, size = 0.9) +
  geom_line(data = region_curves_kin, aes(dist_as, fitted, color = region_class), linewidth = 1.1) +
  geom_text(data = ann_decay_kin, aes(x = 1, y = y, label = lbl, color = region_class),
            inherit.aes = FALSE, hjust = 0, size = 5.3, parse = TRUE, show.legend = FALSE) +
  scale_color_manual(values = region_colors, labels = region_labels_ss) +
  facet_wrap(~kinase, ncol = 5, labeller = lab_kinase) +
  coord_cartesian(ylim = c(0, 1.81), xlim = c(0, 35)) +
  labs(x = "Distance to active site (Å)",
       y = expression("Per residue median |"*Delta*Delta*G[a]*"|"), color = "Region") +
  theme_classic(base_size = 12) + theme(strip.text = element_text(face = "bold", size = 13.5), axis.text = element_text(size = 10.6), legend.text = element_text(size = 11.1), legend.title = element_text(size = 13.5))
ggsave(paste0(output_dir_figS, "FigS6B_decay_by_region", format_img), p_region_kin, width = 15, height = 3.5)
```


# Decay by direction

## Activating mutations
```{r}
all_ddg_dist_stab<-all_ddg_dist[`mean_kcal/mol`<0]

all_ddg_dist$kinase<-factor(all_ddg_dist$kinase,levels=kinases)
kinase_list <- kinases
results <- fit_multiple_kinases(kinase_list,all_ddg_dist_stab, ycol = "abs_ddg")

stab_all_predicted_data <- rbindlist(results$data)
```
## Inactivating mutations
```{r}
all_ddg_dist_destab<-all_ddg_dist[`mean_kcal/mol`>0]

all_ddg_dist$kinase<-factor(all_ddg_dist$kinase,levels=kinases)
kinase_list <- kinases
results <- fit_multiple_kinases(kinase_list,all_ddg_dist_destab, ycol = "abs_ddg")

destab_all_predicted_data <- rbindlist(results$data)
```
## Per-residue median, activating vs inactivating
```{r}
fit_exp_nls <- function(dt) {
  rss <- function(p, x, y) sum((y - p[1] * exp(p[2] * x))^2)
  opt <- optim(c(a = 1.5, b = -0.05), rss, x = dt$active_site_distance, y = dt$abs_ddg)$par
  nls(abs_ddg ~ a * exp(b * active_site_distance), data = dt,
      start = list(a = opt[1], b = opt[2]))
}

fit_two_directions_for_kinase <- function(kinase_name, data_combined, data_all,
                                          assay_type = "Activity",
                                          point_size = 1,
                                          ylim_val = c(0, 1.6),
                                          anno_size = 3.88) {
  act   <- data_combined[kinase == kinase_name & assay == assay_type & direction == "activating"]
  inact <- data_combined[kinase == kinase_name & assay == assay_type & direction == "inactivating"]
  all   <- data_all[kinase == kinase_name & assay == assay_type]
  if (nrow(act) < 5 || nrow(inact) < 5 || nrow(all) < 5) return(NULL)

  fit_act   <- fit_exp_nls(act)
  fit_inact <- fit_exp_nls(inact)
  fit_all   <- fit_exp_nls(all)

  x_seq       <- seq(3, 35, length.out = 100)
  newx        <- data.frame(active_site_distance = x_seq)
  curve_act   <- data.frame(x = x_seq, y = predict(fit_act,   newdata = newx))
  curve_inact <- data.frame(x = x_seq, y = predict(fit_inact, newdata = newx))
  curve_all   <- data.frame(x = x_seq, y = predict(fit_all,   newdata = newx))

  pa  <- coef(fit_act)
  pi  <- coef(fit_inact)
  pal <- coef(fit_all)

  dhalf <- function(b) round(log(2) / abs(unname(b)), 1)
  k_act    <- bquote(k == .(round(unname(pa["b"]),  3)))
  d_act    <- bquote(d[1/2] == .(dhalf(pa["b"]))  * " Å")
  k_inact  <- bquote(k == .(round(unname(pi["b"]),  3)))
  d_inact  <- bquote(d[1/2] == .(dhalf(pi["b"]))  * " Å")
  anno_all <- bquote(k == .(round(unname(pal["b"]), 3)))

  anno_gap     <- ylim_val[2] * 0.08
  anno_y_inact <- ylim_val[2] * 0.97
  anno_y_act   <- ylim_val[2] * 0.72

  err_layers <- if ("err" %in% names(act)) {
    list(
      geom_linerange(data = inact,
                     aes(x = active_site_distance,
                         ymin = pmax(0, abs_ddg - err),
                         ymax = abs_ddg + err),
                     color = "salmon", alpha = 0.5, linewidth = 0.3),
      geom_linerange(data = act,
                     aes(x = active_site_distance,
                         ymin = pmax(0, abs_ddg - err),
                         ymax = abs_ddg + err),
                     color = "cornflowerblue", alpha = 0.5, linewidth = 0.3)
    )
  } else NULL

  ggplot() +
    err_layers +
    geom_point(data = inact, aes(x = active_site_distance, y = abs_ddg), color = "salmon", alpha = 0.7, size = point_size) +
    geom_point(data = act,   aes(x = active_site_distance, y = abs_ddg), color = "cornflowerblue",  alpha = 0.7, size = point_size) +
    geom_line(data = curve_act,   aes(x = x, y = y), color = "blue",  linewidth = 1,linetype = "dashed") +
    geom_line(data = curve_inact, aes(x = x, y = y), color = "red",   linewidth = 1,linetype = "dashed") +
    annotate("text", x = 25, y = anno_y_act,             label = paste(deparse(k_act),   collapse = ""), color = "blue", parse = TRUE, size = anno_size) +
    annotate("text", x = 25, y = anno_y_act - anno_gap,  label = paste(deparse(d_act),   collapse = ""), color = "blue", parse = TRUE, size = anno_size) +
    annotate("text", x = 25, y = anno_y_inact,            label = paste(deparse(k_inact), collapse = ""), color = "red",  parse = TRUE, size = anno_size) +
    annotate("text", x = 25, y = anno_y_inact - anno_gap, label = paste(deparse(d_inact), collapse = ""), color = "red",  parse = TRUE, size = anno_size) +

    coord_cartesian(ylim = ylim_val, xlim = c(3, 35)) +
    labs(x = "Distance to the active site (Å)",
         y = expression("|"*Delta*Delta*G[a]*"| (kcal/mol)"),
         title = relabel_jnk2(kinase_name)) +
    scale_x_continuous(breaks = seq(5, 35, by = 5)) +
    theme_classic(base_size = 13) +
    theme(legend.position = "none",
          axis.text  = element_text(size = 10.9))
}

per_residue_act_split <- all_ddg_dist[`mean_kcal/mol` < 0, .(
  abs_ddg      = median(abs_ddg, na.rm = TRUE),
  err          = 1.2533 * sd(abs_ddg, na.rm = TRUE) / sqrt(sum(!is.na(abs_ddg))),
  active_site_distance = active_site_distance[1],
  direction    = "activating"
), by = .(Pos, kinase, assay, Pos_al)]

per_residue_inact_split <- all_ddg_dist[`mean_kcal/mol` > 0, .(
  abs_ddg      = median(abs_ddg, na.rm = TRUE),
  err          = 1.2533 * sd(abs_ddg, na.rm = TRUE) / sqrt(sum(!is.na(abs_ddg))),
  active_site_distance = active_site_distance[1],
  direction    = "inactivating"
), by = .(Pos, kinase, assay, Pos_al)]

per_residue_split <- rbind(per_residue_act_split, per_residue_inact_split)
per_residue_split[, kinase := factor(kinase, levels = kinases)]

plots_split <- lapply(kinases, fit_two_directions_for_kinase,
                      data_combined = per_residue_split,
                      data_all      = per_residue_ddg,
                      anno_size     = 4.8)
names(plots_split) <- kinases
plots_split <- lapply(plots_split, function(p)
  if (is.null(p)) NULL else p + labs(y = expression("Per residue median |"*Delta*Delta*G[a]*"|")))

p_exp_decay_split <- wrap_plots(plots_split, ncol = 5)

ggsave(p_exp_decay_split,
       filename = paste0(output_dir_fig, "Fig4B_decay_split", format_img),
       height = 3.5, width = 17, limitsize = FALSE)
```

## Inhibitory d1/2 with 95% CI
```{r}
dhalf_ci <- rbindlist(lapply(kinases, function(k) {
  dt <- per_residue_inact_split[kinase == k & assay == "Activity" &
                                is.finite(active_site_distance) & is.finite(abs_ddg)]
  fit <- tryCatch(fit_exp_nls(dt), error = function(e) NULL)
  if (is.null(fit)) return(data.table(kinase = k, n_pos = nrow(dt), k = NA_real_,
                                      d_half = NA_real_, ci_lo = NA_real_, ci_hi = NA_real_, r2 = NA_real_))
  b  <- unname(coef(fit)["b"])
  se <- tryCatch(summary(fit)$coefficients["b", "Std. Error"], error = function(e) NA_real_)
  ci_lo <- if (is.finite(se)) log(2) / (abs(b) + 1.96 * se) else NA_real_
  ci_hi <- if (is.finite(se)) log(2) / (abs(b) - 1.96 * se) else NA_real_
  data.table(kinase = k, n_pos = nrow(dt), k = round(b, 4),
             d_half = round(log(2) / abs(b), 1),
             ci_lo = round(ci_lo, 1), ci_hi = round(ci_hi, 1),
             r2 = round(cor(predict(fit), dt$abs_ddg)^2, 3))
}))
dhalf_ci[, kinase := factor(kinase, levels = kinases)]
setorder(dhalf_ci, kinase)
print(dhalf_ci)
message(sprintf("Mean inhibitory d1/2 across %d kinases = %.1f A", nrow(dhalf_ci), mean(dhalf_ci$d_half, na.rm = TRUE)))
cat(sprintf("Variance explained (R2), inhibitory fits: mean %.0f%% (range %.0f%% to %.0f%%)\n",
            100 * mean(dhalf_ci$r2), 100 * min(dhalf_ci$r2), 100 * max(dhalf_ci$r2)))

dhalf_pooled <- mean(dhalf_ci$d_half, na.rm = TRUE)
cat(sprintf("Pooled inhibitory d1/2 = %.1f A, shell spacing = %.1f A -> %.1f%% reduction in |ddGa| per shell\n",
            dhalf_pooled, shell_spacing, 100 * (1 - 2^(-shell_spacing / dhalf_pooled))))
```

# All mutations, density view
```{r}
decay_density_panel <- function(k, ylim_val = c(0, 2.5)) {
  d <- all_ddg_dist[kinase == k & assay == "Activity" &
                    is.finite(active_site_distance) & is.finite(abs_ddg)]
  if (nrow(d) < 5) return(NULL)
  fit   <- fit_exp_nls(d)
  p     <- coef(fit)
  x_seq <- seq(3, 35, length.out = 100)
  curve <- data.frame(x = x_seq,
                      y = predict(fit, newdata = data.frame(active_site_distance = x_seq)))
  d_half     <- log(2) / abs(unname(p["b"]))
  a0_anno    <- bquote("|"*Delta*Delta*"Ga|"[0] == .(round(unname(p["a"]), 3)))
  k_anno     <- bquote(k == .(round(unname(p["b"]), 3)))
  dhalf_anno <- bquote(d[1/2] == .(round(d_half, 1)) * " Å")

  ggplot(d, aes(x = active_site_distance, y = abs_ddg)) +
    geom_pointdensity(size = 0.4, adjust = 1) +
    scale_color_gradient(low = "lightgrey", high = "black", name = "Local\ndensity") +
    geom_line(data = curve, aes(x = x, y = y), color = "black", linewidth = 1, inherit.aes = FALSE) +
    annotate("text", x = 26, y = ylim_val[2] * 0.94, label = paste(deparse(a0_anno),    collapse = ""), parse = TRUE, size = 4.9) +
    annotate("text", x = 26, y = ylim_val[2] * 0.86, label = paste(deparse(k_anno),     collapse = ""), parse = TRUE, size = 4.9) +
    annotate("text", x = 26, y = ylim_val[2] * 0.78, label = paste(deparse(dhalf_anno), collapse = ""), parse = TRUE, size = 4.9) +
    coord_cartesian(ylim = ylim_val, xlim = c(3, 35)) +
    scale_x_continuous(breaks = seq(5, 35, by = 5)) +
    labs(x = "Distance to the active site (Å)",
         y = expression("|"*Delta*Delta*G[a]*"| (kcal/mol)"),
         title = relabel_jnk2(k)) +
    theme_classic(base_size = 13) +
    theme(legend.position = "none", axis.text = element_text(size = 11.4))
}

plots_density        <- Filter(Negate(is.null), lapply(kinases, decay_density_panel))
p_exp_decay_density  <- wrap_plots(plots_density, ncol = 5)

ggsave(p_exp_decay_density,
       filename = paste0(output_dir_figS, "FigS6A_decay_all_mutations", format_img),
       height = 3.5, width = 18, limitsize = FALSE)
```

# ChimeraX

## Median |ddGa|, all mutations
```{r}
all_predicted_data <- all_predicted_data[, !duplicated(names(all_predicted_data)), with = FALSE]

all_ddg_dist_median_abs_pos <- all_predicted_data[, .(median_abs_pos = median(abs_ddg, na.rm = TRUE)),by = .(Pos, kinase, assay,Pos_al)]

all_predicted_data<-merge(all_predicted_data,all_ddg_dist_median_abs_pos,by = c("Pos", "kinase", "assay","Pos_al"),
                  all.x = TRUE)
unique_kinases <- unique(all_predicted_data$kinase)

all_ddg_dist_chim<-unique(all_predicted_data[,.(kinase,median_abs_pos,Pos)])
for (k in unique_kinases) {
  sig_residues <- all_ddg_dist_chim[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: MedianAbsPos",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$median_abs_pos)
  )

  file_name <- paste0(output_dir_chimera, "median_abs_pos_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```

## Median |ddGa|, activating mutations

```{r}
stab_all_predicted_data <- stab_all_predicted_data[, !duplicated(names(stab_all_predicted_data)), with = FALSE]

stab_all_ddg_dist_median_abs_pos <- stab_all_predicted_data[, .(median_abs_pos = median(abs_ddg, na.rm = TRUE)),
                                by = .(Pos, kinase, assay,Pos_al)]

stab_all_predicted_data<-merge(stab_all_predicted_data,stab_all_ddg_dist_median_abs_pos,by = c("Pos", "kinase", "assay","Pos_al"),
                  all.x = TRUE)

unique_kinases <- unique(all_ddg_dist$kinase)
stab_all_predicted_data<-unique(stab_all_predicted_data[,.(kinase,median_abs_pos,Pos)])
for (k in unique_kinases) {
  sig_residues <- stab_all_predicted_data[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: StabMedianAbsPos",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$median_abs_pos)
  )

  file_name <- paste0(output_dir_chimera, "stab_median_abs_pos_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```
## Median |ddGa|, inactivating mutations

```{r}
destab_all_predicted_data <- destab_all_predicted_data[, !duplicated(names(destab_all_predicted_data)), with = FALSE]

destab_all_ddg_dist_median_abs_pos <- destab_all_predicted_data[, .(median_abs_pos = median(abs_ddg, na.rm = TRUE)),
                                by = .(Pos, kinase, assay,Pos_al)]

destab_all_predicted_data<-merge(destab_all_predicted_data,destab_all_ddg_dist_median_abs_pos,by = c("Pos", "kinase", "assay","Pos_al"),
                  all.x = TRUE)

unique_kinases <- unique(all_ddg_dist$kinase)
destab_all_predicted_data<-unique(destab_all_predicted_data[,.(kinase,median_abs_pos,Pos)])
for (k in unique_kinases) {
  sig_residues <- destab_all_predicted_data[kinase == k]

  if (nrow(sig_residues) == 0) next

  chimerax_lines <- c(
    "attribute: DestabMedianAbsPos",
    "recipient: residues",
    paste0("\t:",sig_residues$Pos, "\t",sig_residues$median_abs_pos)
  )

  file_name <- paste0(output_dir_chimera, "destab_median_abs_pos_", k, ".defattr")

  writeLines(chimerax_lines, file_name)
}
```


# Save predicted data

```{r}
fwrite(all_predicted_data,file=paste0(results_path,"all_predicted_data.csv"))
```

# ChimeraX: active-site distance (SRC)
```{r}
src_dist <- unique(all_ddg_dist[kinase == "SRC", .(Pos, active_site_distance)])
src_dist <- src_dist[!is.na(active_site_distance)]

chimerax_lines <- c(
  "attribute: activesitedistance",
  "recipient: residues",
  paste0("\t:", src_dist$Pos, "\t", src_dist$active_site_distance)
)

writeLines(chimerax_lines, paste0(output_dir_chimera, "active_site_distance_SRC.defattr"))
```


